Disentangling patterns of gene expression in high grade serous ovarian cancer by matrix factorization

### 2020-08-17 MSc_Dissertation_B156476.tex
Post submission: fixed typo in abstract; added my name.  Will now make repository public

\begin{mdframed}
\begin{center}
\large
\vspace{0.2in}
Student Exam Number: \bf{B156476}
\vspace{0.2in}
\end{center}
\end{mdframed}

@@

\begin{abstract}
Gene expression analysis -- transcriptomics -- of a set biopsied tumour tissues can identify genes whose expression is linked to the pathological processes of the tumour.  In contrast to supervised contrastive approaches unsupervised approaches work with datasets of mixed and unknown tumour status.   Matrix factorization (MF) methods have become a popular means of reducing and ``disentangling'' the salient components of variation in such datasets, known as metagenes.  From these metagenes the most influential genes can be identified, and when subjected to gene enrichment analysis against the Gene Ontology (GO) can provide insights into the underlying biological processes in play.

In this work the biological target is high-grade serous ovarian cancer (HGSOC) -- the most common and aggressive type of ovarian cancer.   Two gene expression datasets are analysed; the Austrailian Ovarian Cancer Study dataset (N=80) has patient survival and genomic metadata, while the larger dataset from The Cancer Genome Atlas (N=374) has only survival metadata.  This motivates an approach of unsupervised metagenes discovered by MF techniques in the larger dataset then transferring that information to the smaller dataset to test for their power to stratify patient survival and for correlation with genomic features.

>>>>

\begin{mdframed}
\begin{center}
\large
\vspace{0.2in}
Ian Poole \\
Student Examination Number: \bf{B156476}
\vspace{0.2in}
\end{center}
\end{mdframed}

@@

\begin{abstract}
Gene expression analysis -- transcriptomics -- of a set biopsied tumour tissues can identify genes whose expression is linked to the pathological processes of the tumour.  In contrast to supervised contrastive approaches unsupervised techniques work with datasets of mixed and unknown tumour status.   Matrix factorization (MF) methods have become a popular means of reducing and ``disentangling'' the salient components of variation in such datasets, known as metagenes.  From these metagenes the most influential genes can be identified, and when subjected to gene enrichment analysis against the Gene Ontology (GO) can provide insights into the underlying biological processes in play.

In this work the biological target is high-grade serous ovarian cancer (HGSOC) -- the most common and aggressive type of ovarian cancer.   Two gene expression datasets are analysed; the Australian Ovarian Cancer Study dataset (N=80) has patient survival and genomic metadata, while the larger dataset from The Cancer Genome Atlas (N=374) has only survival metadata.  This motivates an approach of unsupervised metagenes discovered by MF techniques in the larger dataset then transferring that information to the smaller dataset to test for their power to stratify patient survival and for correlation with genomic features.
